An Abstraction-based Approach to Network Synthesis

### 2016-11-01 paper/abstraction/paper.tex
pass on 2

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/example}
  \caption{An example data center network.}
  \label{fig:example}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/example2}
  \caption{A modified version of the example in Figure~\ref{fig:example}.}
  \label{fig:example2}
  \vspace{-1em}
\end{figure}

We motivate \sysname using an example data center network adapted from prior work~\cite{propane}. Our network, shown in Figure~\ref{fig:example}, is a standard folded-Clos topology with four pods. It has eight top-of-rack (ToR) switches, T[1--8]; eight aggregation switches, A[1--8]; and two spine switches, S[1--2]. Groups of addresses, P[1--8], are for services hosted in each rack. Such groups are called address {\em prefixes} because they are compactly encoded using the common prefix that addresses share.

The figure also shows the intended policy. Internal connectivity should be complete, i.e., all nodes must be able to talk to each other. Pods 1 and 2 host services that are meant to be globally accessible and thus their addresses should be announced externally, but only after the individual prefixes P[1--4] are aggregated into a single prefix PG. Finally, pods 3 and 4 host services that are not meant to be externally visible and thus their address should not be announced externally.

To correctly configure the policy above, the network engineer's task is to generate low-level configuration commands for each switch in the network. This implies making sure, for instance, that all switch interfaces are assigned correct addresses; all routing adjacencies are correctly configured (e.g., T1's configuration includes A1 as neighbor and vice-versa); all switches announce their own addresses (so that they can be reached for management purposes) and the ToRs announce the correct prefixes for their services; all switches  forward the prefix announcements that they should to each neighbor and not forward those that they should not (e.g., the spines should forward prefixes for local services to internal neighbors but not external neighbors); and the spines announce externally only the aggregate global prefix. As many have observed before, such configuration tasks are highly challenging and, consequently, configuration errors are the leading contributor to network outages~\cite{x,y,z,batfish,propane}.

\subsection{Challenges with today's template systems}

To reduce errors, the network engineering community is moving toward a template-based approach. Here, instead of authoring one configuration file per router, they author one configuration template per {\em role}, where each role is played by one or more routers.\footnote{If a role is played by devices from different vendors, there is one template per role-vendor combination because templates use vendor configuration language.} For instance, the network in Figure~\ref{fig:example} has five roles: spine, global aggregator, global ToR, local aggregator, and local ToR. Role templates have parameters for various aspects of the configuration (e.g., interface addresses, list of neighbors), and they are compiled to device configurations by instantiating the parameters using a database that stores parameter values for each device.

Templates provide two key advantages. First, they help eliminate a class of configuration errors that stem from typos and inconsistencies, e.g., using the wrong address for a switch, having inconsistent addresses for the two ends of a link, and failing to configure one side of a routing adjacency.  Second, templates have the potential for re-use when the topology changes while maintaining its basic structure (e.g., adding a new pod to the data center) or multiple networks with similar structures need to be instantiated (e.g., in an organization that runs multiple data centers).

%Second, the engineers need to author and reason about fewer independent ``files"---18 versus 5 in our example but the ratio can be multiple orders of magnitude for large networks.

Today's template systems however, are far from perfect. While they reduce a class of errors, they do not help with deeper, logical bugs.
%, e.g., incorrect strategy for determining how announcements to a neighbor are forwarded
%, not realizing when or where aggregation should be used,
%and not understanding the interaction between configuration commands.
To author correct templates, engineers must still do themselves the conceptually challenging task of breaking down complex, global policies into role templates such that the collective behavior of the templates (upon compilation to configurations) correctly implements the policy under all failure conditions.

As a result, the types of configuration design errors in our example network pointed in prior work~\cite{propane} can occur in template design as well. For instance, to keep local prefixes internal the engineer may implement announcement forwarding on spines based on the identity of the neighbor, i.e., do not announce externally anything heard from local aggregators. This implementation works correctly when there are no failures but local prefixes can leak to the outside world when failures disconnect a spine from both aggregators in a local pod. Specifically, when S1 is disconnected from A5 and A6, it may hear about P5 and P6 from A3 and A4 (via A[5--6]->S2->A[3--4]), and it will then announce those prefixes outside (since they no not come from a local aggregator), leading to a policy violation. A correct implementation here would be for each spine to not advertise externally announcements from local aggregators {\em and} reject announcements that have traversed the other spine (i.e., disallow "valleys," or up-down-up paths).

Lacking correctness guarantees, engineers use simulations to analyze if templates are correct under a range of failure cases (e.g., combinations of two-link failures), but such analysis is highly expensive for large networks and still does not provide strong guarantees beyond the tested cases.

@@

\subsection{Challenges with concrete synthesis}

As an alternative approach to reduce configuration errors, the research community has developed many systems that take network topology and policy as input and output device configurations that are  guaranteed to be correct under all possible failures. Thus, they save the engineers from performing the complex task of breaking down global policies into device configurations.

However, they have two shortcomings, both stemming from operating over concrete topologies. The first is a mismatch in abstraction. Engineers that run large networks tend to think of them, and develop tools for them, in terms of roles and not individual devices. We have thus found that they are reluctant to use a system that does not understand roles as a first-order abstraction. There is no guarantee in these systems that the configurations of two devices that serve the same role will be "similar."  If the engineers wanted to sanity-check or (familiarize themselves with) their output, they'll have to consider all device configurations;  there can be orders of magnitude more devices than roles in the network.

A second shortcoming is lack of support for incremental growth. When engineers want to expand a network, they would have to input the new, planned topology to the system. The new configurations for existing devices may undergo unnecessary changes, beyond those needed to accommodate the expansion. For instance, suppose the engineers want to add another pod with local services to the network in Figure~\ref{fig:example}, and (like current aggregators) the aggregators in the new pod will connect to both spines. In this network, Propane, a concrete synthesis tool, uses prefix lists at spines to ensure that local prefixes are not advertised externally. To accommodate the expansion, the new spine configurations outputted by Propane will not only enable the interfaces that connect to the new aggregators but will also have to update the prefix list filters. A system that supports incremental growth should need only interface configuration changes (not filter changes). It can configure all local ToRs, including the new ones, to attach a tag to their announcements and filter external announcements at the spines based on this tag. Engineers prefer fewer configuration changes because each change runs the risks of poor transient behavior (which synthesis tools do not reason about).

>>>>

\begin{figure}[t!]
  \centering
  \includegraphics[width=2.5in]{figures/example}
  \caption{An example data center network.}
  \label{fig:example}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \centering
  \includegraphics[width=2.5in]{figures/example2}
  \caption{A modified version of the example in Figure~\ref{fig:example}.}
  \label{fig:example2}
  \vspace{-1em}
\end{figure}

We motivate \sysname using an example data center network adapted from prior work~\cite{propane}. Our network, shown in Figure~\ref{fig:example}, is a standard folded-Clos topology with four pods. T[1--8] denote top-of-rack (ToR) switches, A[1--8] denote aggregation switches, and S[1--2] denote spine switches. P[1--8] denote groups of addresses for services hosted in each rack. Such groups are called address {\em prefixes} because they are encoded using the adresses' shared prefix. Assume that all nodes are running the border gateway protocol (BGP), as is common in large data centers~\cite{x}. \todo{more detail on BGP?}

The intended policy for this network is as follows. Internal connectivity should be complete, i.e., all nodes must be able to talk to each other. Services in Pods 1 and 2 should be globally accessible and thus their addresses should be announced externally, but only after aggregating individual prefixes P[1--4] into a single prefix PG. Finally, services in Pods 3 and 4 should be not be externally visible and thus their address should not be announced externally.

To correctly configure the policy above, the network engineer's task is to generate low-level configuration commands for each switch in the network. This implies ensuring, for instance, that all switch interfaces are assigned correct addresses; all routing adjacencies are correctly configured (e.g., T1's configuration includes A1 as neighbor and vice-versa); all switches announce their own addresses (so that they can be reached for management purposes) and the ToRs announce the correct prefixes for their services; all switches  forward the prefix announcements that they should to each neighbor and not forward those that they should not (e.g., the spines should forward prefixes for local services to internal neighbors but not external neighbors); and the spines announce externally only the aggregate global prefix.

As many have observed before, such configuration tasks are highly challenging. Consequently, configuration errors are the leading cause of network outages~\cite{x,y,z,batfish,propane}.

\subsection{Limitations of concrete synthesis}

In response, researchers have developed systems that synthesize low-level device configurations from high-level policy statements~\cite{x,y,propane}.
%They take network topology and policy as input and output device configurations that are  guaranteed to be correct under all possible failures. Thus, they save the engineers from performing the complex task of breaking down global policies into device configurations.
These systems offer appealing guarantees that the generated configurations correctly implement policy, even under arbitrary combinations of failures.

However, our conversations with network engineers reveal that they have two limitations that discourage adoption.  The first is a mismatch in abstraction. Engineers of large networks tend to think of the network, and develop tools, in terms of roles and not individual devices. A {\em role} refers to specific functionality and is played by one or more routers. The network in Figure~\ref{fig:example} has five roles: spine, global aggregator, global ToR, local aggregator, and local ToR.  Engineers are reluctant to use a system that does not understand roles as a first-order abstraction. A superficial use of tags to group similar devices is not enough. For instance, there is no guarantee in concrete synthesis systems that the configurations of two devices with the same role will be similar.  If the engineers wanted to debug, or simply understand, their output, they'll have to consider all device configurations instead of role configurations;  there can be orders of magnitude more devices than roles in the network.

A second limitation is inability to support network evolution. Networks evolve all the time as nodes and links are added or removed and the addresses of services are changed. A system that supports evolution should ensure that correctly supporting such changes requires only small changes to network configurations. Engineers want to minimize changes to device configurations because each change runs the risk of poor, dynamic behavior during the change.

Concrete synthesis tools do not accommodate evolution well. When engineers want to modify the network, they would use the new topology as input to the system. (Old inputs or configurations are not supplied.) In the resulting output, existing device configurations may undergo unnecessary changes, beyond those needed to accommodate the modification. Suppose we want to modify the prefix of local ToR T8 in  Figure~\ref{fig:example}. To accommodate the modification, Propane, a concrete synthesis tool, will not only change T8's configuration but also those of the two spines because it uses prefix lists at spines to ensure that local prefixes are not advertised externally.  As we show later, there is a way to design the configurations of this network in a way that does not require spine configuration changes to accommodate ToR prefix modifications, by exposing the inherent structure in this topology to the synthesis tool.

%nof the new spine configurations outputted by Propane will not only enable the interfaces that connect to the new aggregators but will also have to update the prefix list filters. A system that supports incremental growth should need only interface configuration changes (not filter changes). It can configure all local ToRs, including the new ones, to attach a tag to their announcements and filter external announcements at the spines based on this tag. Engineers prefer fewer configuration changes because each change runs the risks of poor transient behavior (which synthesis tools do not reason about).

%For instance, suppose the engineers want to add another pod with local services to the network in Figure~\ref{fig:example}. In this network, Propane, a concrete synthesis tool, uses prefix lists at spines to ensure that local prefixes are not advertised externally. To accommodate the expansion, the new spine configurations outputted by Propane will not only enable the interfaces that connect to the new aggregators but will also have to update the prefix list filters. A system that supports incremental growth should need only interface configuration changes (not filter changes). It can configure all local ToRs, including the new ones, to attach a tag to their announcements and filter external announcements at the spines based on this tag. Engineers prefer fewer configuration changes because each change runs the risks of poor transient behavior (which synthesis tools do not reason about).

\subsection{Limitations of today's template systems}

To simplify network configuration, and thus reduce errors, the network engineering community is adopting a template-based approach~\cite{x,y}. Instead of authoring one configuration file per router, engineers author one configuration template per role.\footnote{If a role is played by devices from different vendors, there is one template per role-vendor combination because templates use vendor configuration language.} Templates have parameters for various aspects of the configuration (e.g., interface addresses, neighbor list) and are compiled to device configurations by instantiating the parameters using a database of per-device values.

%Templates provide two key advantages. First, they help eliminate a class of configuration errors that stem from typos and inconsistencies, e.g., using the wrong address for a switch, having inconsistent addresses for the two ends of a link, and failing to configure one side of a routing adjacency.  Second, templates have the potential for re-use when the topology changes while maintaining its basic structure (e.g., adding a new pod to the data center) or multiple networks with similar structures need to be instantiated (e.g., in an organization that runs multiple data centers).

While templates operate at the right level of abstraction, today's template systems have two limitations. The first is that they offer no guarantees on correctness. To author correct templates, engineers must do manually the conceptually challenging task of breaking down complex, global policies into role templates such that the collective behavior of the templates (upon compilation to configurations) correctly implements the policy under all failure conditions.

As a result, the types of configuration design errors pointed in prior work~\cite{propane} can occur in template design as well. For instance, to keep local prefixes internal the engineer may implement announcement forwarding on spines based on the identity of the neighbor, i.e., do not announce externally anything heard from local aggregators. This implementation is appealing because it does not require changes to spine configurations when ToR prefixes change. But it works correctly only when there are no failures; local prefixes can leak externally when failures disconnect a spine from both aggregators in a local pod. Specifically, when S1 is disconnected from A5 and A6, it may hear about P5 and P6 from A3 and A4 (via A[5--6]->S2->A[3--4]), and it will then announce those prefixes outside (since they no not come from a local aggregator), leading to a policy violation.

%A correct implementation here would be for each spine to not advertise externally announcements from local aggregators {\em and} reject announcements that have traversed the other spine (i.e., disallow "valleys," or up-down-up paths).

%Lacking correctness guarantees, engineers use simulations to analyze if templates are correct under a range of failure cases (e.g., combinations of two-link failures), but such analysis is highly expensive for large networks and still does not provide strong guarantees beyond the tested cases.

The second limitation of today's template systems is that, like concrete synthesis tools, they do not support network evolution. In theory, templates are appealing because operators do not have to redesign templates when the topology changes while keeping its structure of roles and their inter-connectivity. In practice, however, there is no way to \todo{ ... continue here ...}

@@
